Improving the Replicability of Psychological Science Through Pedagogy

### 2017-06-28 paper/paper.Rmd
Small edits + spellcheck

In the current article, we address the practical barriers of completing replcations as part of required coursework and discuss the methodological decisions that replication projects must address more braodly. Towards this aim, we report the results of replication projects conducted in a graduate-level experimental methods course. Students conducted replications of published articles from the 2015 volume of the journal *Psychological Science*. These studies provide insight into both the difficulties of pedagogical replications and their promise as a method for improving the robustness of psychological research.

@@

To encourage others to share our materials, our course outline, project templates, and assignments are available publicly at [https://osf.io/98ta4/files/](https://osf.io/98ta4/files/). In addition, all of our most recent lecture slides and materials are avaialable on our course website at [http://psych254.stanford.edu](). We hope that others will share and reuse our materials as they consider how best to design a course customized to their context.

@@

In addition to as-needed guidance on projects, students advanced through a formal review process in which each student's work was inspected and critiqued several times by both the instructor and the teaching assistants. After selecting their replication targets, students' first milestone was to fill in the methods section of the @osc2015 replication report template, which was reviewed for feasability and concreteness. This section included a power analysis, proposed sample size, and explicit description of differences from the original study (see course materials for complete template).

@@

Project selection & \multicolumn{1}{p{2in}}{\raggedright Single class replication or small curated list} & \multicolumn{1}{p{2in}}{\raggedright Curated list} & \multicolumn{1}{p{2in}}{\raggedright Open project selection*} \\

Data source & \multicolumn{1}{p{2in}}{\raggedright University or classroom participant pool} & \multicolumn{1}{p{2in}}{\raggedright In-person or online convenience sample} & \multicolumn{1}{p{2in}}{\raggedright Targetted sampling of original populations} \\

@@

Dissemination & \multicolumn{1}{p{2in}}{\raggedright If meeting pre-specified quality standard} & \multicolumn{1}{p{2in}}{\raggedright With instructor approval} & \multicolumn{1}{p{2in}}{\raggedright As class default (subject to disucssion)} \\

@@

Over half of our students contacted original authors to request study materials or clafication on methods and/or analysis, but for the purposes of our class, we chose not to require contact with all original authors. We strongly believe methods and materials should be open to allow for replication and productive science [@nosek2015], and hoped to empower students to engage with and build on the published literature directly -- without personal relationships or even personal contact. That is, we chose to do good-faith replications given all information in the original articles and supplemental online materials. Authors were contacted only in cases where there were issues about access to materials or ambiguities in design or analysis. There are benefits of contacting original authors, however. In addition to professional courtesy, authors can provide valuable feedback or guidance (albeit at the cost of some imposition). In the end, it is up to individual instructors to decide what they believe is best for their students.

@@

We chose to use Amazon Mechanical Turk (AMT) as a platform for our replication studies.^[Repeat administration of empirical work on AMT has been raised as an issue in some prior work [e.g., @chandler2014], but tracking participation in specific paradigms is an open challenge. We did not ask participants whether they had participated in similar research previously, as we suspected that asking this sort of question would lead to a large number of inaccurate responses due to failures to distinguish between related experimental paradigms. In addition, half of the AMT population is estimated to change every six months [@stewart2015], and we expected that most of the studies in our sample that had been conducted on AMT had been performed at least a year previously (though see Supplemental Material for discussion of one particular paradigm that has been used more recently).] This choice was made to facilitate the process of recruiting samples large enough to enable replications of between-subjects designs (which typically require large samples) as well as to help students learn about a valuable resource for recruitment whose use requires some specialized knowledge. As much of psychological research is now conducted online, an additional pedagogical benefit of conducting online replications is to build competence in these skills. Even if students’ current research does not use AMT, these replications allow students to build a skillset that can be used throughout their academic careers. Our general class structure does not depend on the use of AMT, however, and the use of this particular online sample should be viewed as limiting the interpretation of some of our results. Other classes many find participant pools or recruitment of community samples more feasible and appropriate.

@@

### Workflow selection

Our explicit class goal was to provide students with experiences navigating the full ecosystem of open-source scientific tools. For classes at different levels, however, instructors will likely have different goals. It may be preferrable for an intermediate course to focus on a single programming tool (e.g., R) and to use other GUI-based software for creating experiments, so as to minimize the learning burden for students. And for more introductory research methods course, instructors must gauge whether the added difficulty of a programming compoent is appropriate for the skills and backgrounds of their students.

@@

Institutional Review Boards (IRBs) are idiosyncratic across institutions, but we suspect certain ethical concerns will commonly arise when attempting to implement class replication projects. First, instructors may first be concerned that IRB approval is not possible within the time constraints of the class.  To mitigate this issue, we suggest contacting the IRB chair well in advance of the class to determine whether to create a standard protocol that encompassses all replications that will be conducted, or to determine a timeline for submitting individual protocols very early in the course. Both issues may be simplified if the study or pool of studies to be replicated is pre-selected by instructors. Second, review boards may be concerned about risks particular to the collection of data by students, e.g., lack of debriefing or greater risk of breach of confidentiality due to novice experimenters. We recommend building training around ethical issues into the syllabus of replication-based classes. This practice is both positive for studetns and mitigates potential IRB concerns via explicit training.

@@

Our statistical findings mask a more optimistic message: In most of our projects, the next step for a motivated experimenter is clear. For example, in some projects we suspect that a followup could find strong evidence for the phenomenon of interest by titrating the particular planned analysis or the difficulty of the stimulus materials. In others, the difficulty appeared to be statistical power, since differences were in the predicted direction and sometimes reached significance in subsidiary analyses, so a followup would likely require a larger sample. And in some, differences of population would mean that followups would require further stimulus or task adaptation. More generally, we believe that our work here underscores the importance of iterated replication for pinpointing empirical effects and refining theories [see e.g., @lewis2016 for an example of this strategy and some discussion].

@@

This kind of cross-class collaboration would require instructors to make class results more broadly discoverable. We can envision a number of possible avenues for this kind of sharing. For example, if some products are shared openly in a repository like the [Open Science Framework](http://osf.io) or [figshare](http://figshare.com) and tagged appropriately, they will be discoverable via search engines. Alternatively, a more structured method for sharing and discovery would be upload to specific replication curation websites like [PsychFileDrawer](http://psychfiledrawer.org) or CurateScience (http://curatescience.org/). Both of these models assume limited coordination across instructors, but more structured collaborations are of course possible. For example, the [Collaborative Replications and Education Project](https://osf.io/wfc6u/) selected a subsample of important and feasible studies for replication and helped provide materials to instructors with the explicit goal of encouraging cross-lab pedagogical replications.

@@

For each project, we present a summary of the finding, the key statistical test, the impressionistic outcomes of the replication, and -- where applicable -- our assessement of the source of differences in results (often composed with the advice of the original authors). Figure 3 shows a side-by-side comparison of the key visualization for each study.

>>>>

In the current article, we address the practical barriers of completing replications as part of required coursework and discuss the methodological decisions that replication projects must address more broadly. Towards this aim, we report the results of replication projects conducted in a graduate-level experimental methods course. Students conducted replications of published articles from the 2015 volume of the journal *Psychological Science*. These studies provide insight into both the difficulties of pedagogical replications and their promise as a method for improving the robustness of psychological research.

@@

To encourage others to share our materials, our course outline, project templates, and assignments are available publicly at [https://osf.io/98ta4/files/](https://osf.io/98ta4/files/). In addition, all of our most recent lecture slides and materials are available on our course website at [http://psych254.stanford.edu](). We hope that others will share and reuse our materials as they consider how best to design a course customized to their context.

@@

In addition to as-needed guidance on projects, students advanced through a formal review process in which each student's work was inspected and critiqued several times by both the instructor and the teaching assistants. After selecting their replication targets, students' first milestone was to fill in the methods section of the @osc2015 replication report template, which was reviewed for feasibility and concreteness. This section included a power analysis, proposed sample size, and explicit description of differences from the original study (see course materials for complete template).

@@

Project selection & \multicolumn{1}{p{2in}}{\raggedright Single class replication or small curated list} & \multicolumn{1}{p{2in}}{\raggedright Curated list} & \multicolumn{1}{p{2in}}{\raggedright Open project selection} \\

Subject population & \multicolumn{1}{p{2in}}{\raggedright University or classroom participant pool} & \multicolumn{1}{p{2in}}{\raggedright In-person or online convenience sample} & \multicolumn{1}{p{2in}}{\raggedright Targeted sampling of original populations} \\

@@

Dissemination & \multicolumn{1}{p{2in}}{\raggedright If meeting pre-specified quality standard} & \multicolumn{1}{p{2in}}{\raggedright With instructor approval} & \multicolumn{1}{p{2in}}{\raggedright As class default (subject to discussion)} \\

@@

@@

We chose to use Amazon Mechanical Turk (AMT) as a platform for our replication studies.^[Repeat administration of empirical work on AMT has been raised as an issue in some prior work [e.g., @chandler2014], but tracking participation in specific paradigms is an open challenge. We did not ask participants whether they had participated in similar research previously, as we suspected that asking this sort of question would lead to a large number of inaccurate responses due to failures to distinguish between related experimental paradigms. In addition, half of the AMT population is estimated to change every six months [@stewart2015], and we expected that most of the studies in our sample that had been conducted on AMT had been performed at least a year previously (though see Supplemental Material for discussion of one particular paradigm that has been used more recently).] We used AMT to facilitate the process of recruiting samples large enough to enable replications of between-subjects designs (which typically require large samples) as well as to help students learn about a valuable resource for recruitment whose use requires some specialized knowledge. As much of psychological research is now conducted online, an additional pedagogical benefit of conducting online replications is to build competence in these skills. Even if students’ current research does not use AMT, these replications allow students to build a skillset that can be used throughout their academic careers. Our general class structure does not depend on the use of AMT, however, and the use of this particular online sample should be viewed as limiting the interpretation of some of our results. Other classes many find participant pools or recruitment of community samples more feasible and appropriate.

### Workflow tools

Our explicit class goal was to provide students with experiences navigating the full ecosystem of open-source scientific tools. For classes at different levels, however, instructors will likely have different goals. It may be preferable for an intermediate course to focus on a single programming tool (e.g., R) and to use other GUI-based software for creating experiments, so as to minimize the learning burden for students. And for more introductory research methods course, instructors must gauge whether the added difficulty of a programming component is appropriate for the skills and backgrounds of their students.

@@

Over half of our students contacted original authors to request study materials or clarification on methods and/or analysis, but for the purposes of our class, we chose not to require contact with all original authors. We strongly believe methods and materials should be open to allow for replication and productive science [@nosek2015], and hoped to empower students to engage with and build on the published literature directly -- without personal relationships or even personal contact. That is, we chose to do good-faith replications given all information in the original articles and supplemental online materials. Authors were contacted only in cases where there were issues about access to materials or ambiguities in design or analysis. There are benefits of contacting original authors, however. In addition to professional courtesy, authors can provide valuable feedback or guidance (albeit at the cost of some imposition). In the end, it is up to individual instructors to decide what they believe is best for their students.

@@

Institutional Review Boards (IRBs) are idiosyncratic across institutions, but we suspect certain ethical concerns will commonly arise when attempting to implement class replication projects. First, instructors may first be concerned that IRB approval is not possible within the time constraints of the class.  To mitigate this issue, we suggest contacting the IRB chair well in advance of the class to determine whether to create a standard protocol that encompasses all replications that will be conducted, or to determine a timeline for submitting individual protocols very early in the course. Both issues may be simplified if the study or pool of studies to be replicated is pre-selected by instructors. Second, review boards may be concerned about risks particular to the collection of data by students, e.g., lack of debriefing or greater risk of breach of confidentiality due to novice experimenters. We recommend building training around ethical issues into the syllabus of replication-based classes. This practice is both positive for students and mitigates potential IRB concerns via explicit training.

@@

However, our statistical findings mask a more optimistic message: In most of our projects, the next step for a motivated experimenter is clear. For example, in some projects we suspect that a followup could find strong evidence for the phenomenon of interest by titrating the particular planned analysis or the difficulty of the stimulus materials. In others, the difficulty appeared to be statistical power, since differences were in the predicted direction and sometimes reached significance in subsidiary analyses, so a followup would likely require a larger sample. And in some, differences of population would mean that followups would require further stimulus or task adaptation. More generally, we believe that our work here underscores the importance of iterated replication for pinpointing empirical effects and refining theories [see e.g., @lewis2016 for an example of this strategy and some discussion].

@@

This kind of cross-class collaboration would require instructors to make class results more broadly discoverable. We can envision a number of possible avenues for this kind of sharing. For example, if some products are shared openly in a repository like the Open Science Framework (osf.io) or figshare (figshare.com) and tagged appropriately, they will be discoverable via search engines. Alternatively, a more structured method for sharing and discovery would be upload to specific replication curation websites like PsychFileDrawer (psychfiledrawer.org) or CurateScience (curatescience.org). Both of these models assume limited coordination across instructors, but more structured collaborations are of course possible. For example, the Collaborative Replications and Education Project (osf.io/wfc6u) selected a subsample of important and feasible studies for replication and helped provide materials to instructors with the explicit goal of encouraging cross-lab pedagogical replications.

@@

For each project, we present a summary of the finding, the key statistical test, the impressionistic outcomes of the replication, and -- where applicable -- our assessment of the source of differences in results (often composed with the advice of the original authors). Figure 3 shows a side-by-side comparison of the key visualization for each study.
